### 2017-11-28 paper/sections/2_parsplice-keyspace-analysis.tex
paper: addresses #5
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%
%We demonstrate these contributions by changing the input to our molecular
%dynamics simulation, which changes the keyspace access patterns. 
%
%\subsection{Structured Access Regimes}
%
%Figures~\ref{fig:keyspace-analysis_size},~\ref{fig:keyspace-analysis_locality} and~\ref{fig:keyspace-analysis_throughput}.
%%shows ``access regimes", or groups of accesses
%%to the same subset of keys:
%\begin{itemize}
%  \item change immediately, different sizes
%  \item monotonically increasing
%  \item random access to a single key
%  \item early keys accessed more
%\end{itemize}
%
%\noindent\textbf{Conclusion}: unique patterns of a real HPC application
%
%\subsection{Cache Size Analysis}
%
%Figure~\ref{fig:keyspace-analysis_cachesize}.
%
%\noindent\textbf{Conclusion}: workload phases (high request rate to a small
%number of keys and then low request rate to a large number keys) need a dynamic
%load balancing policy.
%
%\subsection{Overfitting Policies}
%
%Figures~\ref{fig:keyspace-analysis_cachesize-dynamic} and~\ref{fig:keyspace-analysis_cachesize-caveats}.
%
%\noindent\textbf{Side Idea}: we need to figure out what ParSplice memory is
%sensitive to: max usage or usage over time.\\
%
%\noindent\textbf{Conclusion}: dynamic policies absorb the cost of a high read
%request rate for a 2.5 hour run, but it is infeasible to do this for every
%combination of system setup, job lengths, parsplice parameters.
%
%\subsubsection*{Discussion}
%Why don't we just an LRU cache:
%\begin{itemize}
%  \item we can do better (workload is structured and has locality)
%  \item finding the size of the cache is hard
%  \item hotspots dissipate too quickly
%\end{itemize}

@@

\subsection{Results and Conclusions}
\label{sec:parsplice-keyspace-analysis}
%IT HAS 8K keys!  How did we take these measurements

@@

\subsubsection*{Experimental Setup} All experiments ran on Trinitite, a Cray
XC40 with 32 Intel Haswell 2.3GHz cores per node.  Each node has 128GB of RAM
and our goal is to limit the size of the cache to 3\% of RAM . Note that this
is an addition to the 30GB that ParSplice uses to manage other ranks on the
same node.  The scalability experiment uses 1 splicer, 1 persistent database, 1
cache process, and up to 2 workers. We scale up to 1024 tasks, which spans 32
nodes and disable hyper-threading because we experience unacceptable
variability in performance. For the rest of the experiments, we use 8 nodes, 1
splicer, 1 persistent database, 1 cache process, 1 worker, and up to 256 tasks.
The keyspace analysis that follows is for the cache on the worker node, which
is circled in Figure~\ref{fig:parsplice}.

@@

\subsubsection*{Scalability} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (text annotations) and request load (bars) after a one hour run
with a different number of tasks (\(x\) axis). While the keyspace size and
capacity is modest the memory usage scales linearly with the number
of tasks, which is a problem if we want to scale to Trinitite's 3000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 8GB for one cache.  This memory utilization easily eclipses
the 3\% memory usage per node threshold we set earlier, even without factoring
in the usage from other workers.

\subsubsection*{An active but small keyspace}

@@

\subsubsection*{Entropy increases over time} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) look like
the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the key access patterns and
adapt the storage system accordingly (this motivates
Section~\S\ref{sec:dom-specific}).

>>>>

%\subsubsection*{Discussion}
%Why don't we just an LRU cache:
%\begin{itemize}
%  \item we can do better (workload is structured and has locality)
%  \item finding the size of the cache is hard
%  \item hotspots dissipate too quickly
%\end{itemize}

@@

%IT HAS 8K keys!  How did we take these measurements
\subsection{Experimental Setup}

@@

All experiments ran on Trinitite, a Cray XC40 with 32 Intel Haswell 2.3GHz
cores per node.  Each node has 128GB of RAM and our goal is to limit the size
of the cache to 3\% of RAM . Note that this is an addition to the 30GB that
ParSplice uses to manage other ranks on the same node.  The scalability
experiment uses 1 splicer, 1 persistent database, 1 cache process, and up to 2
workers. We scale up to 1024 tasks, which spans 32 nodes and disable
hyper-threading because we experience unacceptable variability in performance.
For the rest of the experiments, we use 8 nodes, 1 splicer, 1 persistent
database, 1 cache process, 1 worker, and up to 256 tasks.  The keyspace
analysis that follows is for the cache on the worker node, which is circled in
Figure~\ref{fig:parsplice}.

\subsection{Results and Observations}
\label{sec:parsplice-keyspace-analysis}

@@

Our analysis shows that ParSplice accesses keys in a structured and predictable
way. The following 3 observations shape the policies we design later in the
paper.

\subsubsection{\underline{Scalability}} Figure~\ref{fig:methodology-keyspace} shows the
keyspace size (text annotations) and request load (bars) after a one hour run
with a different number of tasks (\(x\) axis). While the keyspace size and
capacity is modest the memory usage scales linearly with the number
of tasks, which is a problem if we want to scale to Trinitite's 3000 cores.
Furthermore, the size of the keyspace also increases linearly with the length
of the run.  Extrapolating these results puts an 8 hour run across all 100
Trinitite nodes at 8GB for one cache.  This memory utilization easily eclipses
the 3\% memory usage per node threshold we set earlier, even without factoring
in the usage from other workers.

\subsubsection{\underline{An active but small keyspace}}

@@

\subsubsection{\underline{Entropy increases over time}} The reads per second in
Figure~\ref{fig:motivation-regimes} show that the number of requests decreases
and the number of active keys increases over time. The resulting key access
imbalance for the two growth rates in Figure~\ref{fig:motivation-regimes} are
shown in Figure~\ref{fig:methodology-keys}, where reads are plotted for each
unique state, or key, along the \(x\) axis. Keys are more popular than others
(up to \(5\times\)) because worker tasks start generating states with different
coordinates later in the run.  Figure~\ref{fig:methodology-keys} shows that the
number of reads changes with different initial conditions (\(\Delta\)), but
that the spatial locality of key accesses is similar ({\it e.g.}, some keys are
still \(5\times\) more popular than others).
Figure~\ref{fig:motivation-regimes} shows how entropy for different growth
rates has temporal locality, as the reads per second for \(\Delta_2\) look like
the reads per second for \(\Delta_1\) stretched out along the time axis.
Trends also exist for temperature and number of workers but are omitted here
for space. This structure means that we can learn the key access patterns and
adapt the storage system accordingly (this motivates
Section~\S\ref{sec:dom-specific}).
